Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-04 paper.qmd
Try to tune with different hyper parameter choices, but nothing much changed with default settings. Therefore, the best with tSNE

```{r}
#| label: read-mnist-nldr
# Read a variety of different NLDR representations of mnist
# and plot them on same aspect ratio
tsne_mnist <- read_rds("data/mnist/mnist_tsne89.rds")

@@

error_plot_mnist <- ggplot(error_mnist, 
                          aes(x = a1, 
                              y = MSE, 
                              colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_x_continuous(breaks = sort(unique(error_mnist$a1))[seq(1, length(unique(error_mnist$a1)), by = 5)]) +
  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a', "#ff7f00",'#a65628')) +
  scale_y_log10() +
  ylab("log(MSE)") +
  xlab(expression(paste("binwidth (", a[1], ")"))) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))

@@

```{r}
#| fig-cap: ""
#| label: fig-mnist-mse
#| fig-pos: H
#| out-height: 100%

@@

pacmap_minst <- read_rds("data/mnist/mnist_pacmap.rds") |> 
  select(PaCMAP1, PaCMAP2, ID)
mnist_scaled_obj <- gen_scaled_data(
  data = pacmap_minst)
pacmap_minst_scaled <- mnist_scaled_obj$scaled_nldr

pacmap_plot_mnist <- pacmap_minst_scaled |> 
  ggplot(aes(x = PaCMAP1, 
             y = PaCMAP2)) + 
  geom_point(alpha=0.1, color = "#000000")

# read the png file from device 
left_top_img <- readPNG("figures/mnist/mnist_digit1_img_lt.png", native = TRUE)
center_img <- readPNG("figures/mnist/mnist_digit1_img_c.png", native = TRUE)
right_bottom_img <- readPNG("figures/mnist/mnist_digit1_img_rb.png", native = TRUE)

pacmap_plot_mnist <- pacmap_plot_mnist +
    inset_element(p = left_top_img, 
                  left = 0.8, 
                  bottom = 0.85, 
                  right = 0, 
                  top = 0.95) +
    inset_element(p = center_img, 
                  left = 1.2, 
                  bottom = 0.6, 
                  right = 0, 
                  top = 0.7) +
    inset_element(p = right_bottom_img, 
                  left = 1.7, 
                  bottom = 0.48, 
                  right = 0, 
                  top = 0.58)
```

@@

pacmap_plot_mnist
```

@@

mnist_model <- fit_highd_model(
  training_data = training_data_mnist,
  emb_df = pacmap_minst_scaled,
  bin1 = num_bins_x_mnist,
  r2 = r2_mnist,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "PC"
)

@@

trimesh_removed_mnist <- ggplot() + 
  geom_segment(data = tr_from_to_df_mnist, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#33a02c") +
  geom_point(data = pacmap_minst_scaled,
             aes(
               x = PaCMAP1,
               y = PaCMAP2
             ),
             alpha=0.5, 
             size = 0.1) +
  coord_fixed() +
  interior_annotation("a", sc_ltr_pos_mnist)

@@

## Compute error
error_df <- augment(
  df_bin_centroids = df_bin_centroids_mnist, 
  df_bin = df_bin_mnist, 
  training_data = training_data_mnist, 
  newdata = NULL, 
  type_NLDR = "PaCMAP", 
  col_start = "PC")

@@

## To join embedding
error_df <- error_df |>
  bind_cols(pacmap_minst_scaled |> 
              select(-ID))

error_plot_mnist <- error_df |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2, 
             color = type, 
             group = ID)) +
  geom_point(alpha=0.5, 
             size = 0.1) +
  scale_colour_brewer(
    type = "qual",
    palette = 2
  ) + 
  theme(
      legend.position = "bottom",
      legend.text = element_text(size = 10),          # Increase legend text size
      legend.key.size = unit(1, "lines")              # Increase size of legend keys
  ) +
  guides(
      color = guide_legend(
          override.aes = list(size = 2.5, alpha = 1)      # Increase point size and opacity in legend
      )
  ) 
```

>>>>

```{r}
#| label: read-mnist-nldr
# Read a variety of different NLDR representations of mnist
# and plot them on same aspect ratio
tsne_mnist <- read_rds("data/mnist/mnist_tsne30.rds")

@@

error_plot_mnist <- ggplot(error_mnist, 
                          aes(x = a1, 
                              y = MSE, 
                              colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_x_continuous(breaks = sort(unique(error_mnist$a1))[append(seq(1, length(unique(error_mnist$a1)), by = 5), 24)]) +
  scale_color_manual(values=c('#e41a1c','#377eb8','#4daf4a', "#ff7f00",'#a65628')) +
  scale_y_log10() +
  ylab("log(MSE)") +
  xlab(expression(paste("binwidth (", a[1], ")"))) +
  theme_minimal() +
  theme(panel.border = element_rect(fill = 'transparent'),
        plot.title = element_text(size = 12, hjust = 0.5, vjust = -0.5),
        axis.ticks.x = element_line(),
        axis.ticks.y = element_line(),
        legend.position = "none",
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7))

@@

```{r}
#| fig-cap: "Different methods with default hyper-parameter settings"
#| label: fig-mnist-mse
#| fig-pos: H
#| out-height: 100%

@@

mnist_scaled_obj <- gen_scaled_data(
  data = tsne_mnist)
tsne_minst_scaled <- mnist_scaled_obj$scaled_nldr

tsne_plot_mnist <- tsne_minst_scaled |> 
  ggplot(aes(x = tSNE1, 
             y = tSNE2)) + 
  geom_point(alpha=0.1, color = "#000000")

# # read the png file from device 
# left_top_img <- readPNG("figures/mnist/mnist_digit1_img_lt.png", native = TRUE)
# center_img <- readPNG("figures/mnist/mnist_digit1_img_c.png", native = TRUE)
# right_bottom_img <- readPNG("figures/mnist/mnist_digit1_img_rb.png", native = TRUE)
# 
# tsne_plot_mnist <- tsne_plot_mnist +
#     inset_element(p = left_top_img, 
#                   left = 0.8, 
#                   bottom = 0.85, 
#                   right = 0, 
#                   top = 0.95) +
#     inset_element(p = center_img, 
#                   left = 1.2, 
#                   bottom = 0.6, 
#                   right = 0, 
#                   top = 0.7) +
#     inset_element(p = right_bottom_img, 
#                   left = 1.7, 
#                   bottom = 0.48, 
#                   right = 0, 
#                   top = 0.58)
```

@@

tsne_plot_mnist
```

@@

mnist_model <- fit_highd_model(
  training_data = training_data_mnist,
  emb_df = tsne_minst_scaled,
  bin1 = num_bins_x_mnist,
  r2 = r2_mnist,
  is_bin_centroid = TRUE,
  is_rm_lwd_hex = FALSE,
  col_start_highd = "PC"
)

@@

trimesh_removed_mnist <- ggplot() + 
  geom_segment(data = tr_from_to_df_mnist, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#33a02c") +
  geom_point(data = tsne_minst_scaled,
             aes(
               x = tSNE1,
               y = tSNE2
             ),
             alpha=0.5, 
             size = 0.1) +
  coord_fixed() +
  interior_annotation("a", sc_ltr_pos_mnist)

@@

## Compute error
error_df <- augment(
  df_bin_centroids = df_bin_centroids_mnist, 
  df_bin = df_bin_mnist, 
  training_data = training_data_mnist, 
  newdata = NULL, 
  type_NLDR = "tSNE", 
  col_start = "PC")

@@

## To join embedding
error_df <- error_df |>
  bind_cols(tsne_minst_scaled |> 
              select(-ID))

error_plot_mnist <- error_df |>
  ggplot(aes(x = tSNE1,
             y = tSNE2, 
             color = type, 
             group = ID)) +
  geom_point(alpha=0.5, 
             size = 0.1) +
  scale_colour_brewer(
    type = "qual",
    palette = 2
  ) + 
  theme(
      legend.position = "bottom",
      legend.text = element_text(size = 10),          # Increase legend text size
      legend.key.size = unit(1, "lines")              # Increase size of legend keys
  ) +
  guides(
      color = guide_legend(
          override.aes = list(size = 2.5, alpha = 1)      # Increase point size and opacity in legend
      )
  ) 
```
